---
output: 
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# schtools

Schloss Lab Tools for Reproducible Microbiome Research 💩

<!-- badges: start -->
[![check](https://github.com/SchlossLab/schtools/workflows/check/badge.svg)](https://github.com/SchlossLab/schtools/actions)
[![codecov](https://codecov.io/gh/SchlossLab/schtools/branch/master/graph/badge.svg)](https://app.codecov.io/gh/SchlossLab/schtools?branch=main)
[![license](https://img.shields.io/badge/license-MIT-blue.svg)](https://github.com/SchlossLab/schtools/blob/main/LICENSE.md)
[![CRAN status](https://www.r-pkg.org/badges/version/schtools)](https://CRAN.R-project.org/package=schtools)
[![Conda](https://img.shields.io/conda/vn/conda-forge/r-schtools)](https://anaconda.org/conda-forge/r-schtools)
![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/schtools)
<!-- badges: end -->

A collection of useful functions and example code created and used by the
Schloss Lab for reproducible microbiome research. Perform common tasks like read
output files from [mothur](https://mothur.org), tidy up your microbiome data,
and format rmarkdown documents for publication. See the
[website](http://www.schlosslab.org/schtools/) for more information,
documentation, and examples.

## Installation

You can install the latest release from [CRAN](https://cran.r-project.org/package=schtools):

```{r, eval = FALSE}
install.packages("schtools")
```

or the development version from [GitHub](https://github.com/SchlossLab) with:

```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("SchlossLab/schtools")
```

## Help & Contributing

If you come across a bug, [open an
issue](https://github.com/SchlossLab/schtools/issues) and include a [minimal
reproducible example](https://www.tidyverse.org/help/).

If you'd like to contribute, see our guidelines
[here](http://www.schlosslab.org/schtools/CONTRIBUTING.html).

## Code of Conduct

Please note that the schtools project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

## License

The schtools package is licensed under 
[the MIT license](https://github.com/SchlossLab/schtools/blob/main/LICENSE.md).
Text and images included in this repository
are licensed under the [CC BY 4.0 license](https://creativecommons.org/licenses/by/4.0/).

## Citation

```{r cite, echo = FALSE, comment = '', warning = FALSE}
citation(package = 'schtools')
```

