---
title: "Introduction to schtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to schtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

```{r deps}
library(schtools)
```

## Parsing mothur output files

### Taxonomy files

mothur formats taxonomy files as tab-separated values (tsv). 
You can use `read_tax()` to parse the taxonomy data and create separate columns
for each taxonomic level.

```{r read_tax}
tax_filepath <- system.file("extdata", "test.taxonomy", package = "schtools")
tax_dat <- read_tax(tax_filepath)
head(tax_dat)
```

The column `label_html` provides html that correctly italicizes the genus name
without italicizing the OTU label. This can be used with
`ggtext::element_markdown()` to make nice plots:

```{r italic-genus}
library(dplyr)
library(ggplot2)
library(ggtext)
library(tidyr)
set.seed(20220427)
pick_random_bool <- function(n = nrow(tax_dat)) {
        sample(c(TRUE, FALSE), n, replace = TRUE)
    }
tax_dat %>% 
    select(label_html) %>% 
    mutate(sample1 = pick_random_bool(),
           sample2 = pick_random_bool(),
           sample3 = pick_random_bool() 
           ) %>% 
    pivot_longer(starts_with('sample'), 
                 names_to = 'sample', values_to = 'presence') %>% 
    ggplot(aes(x = sample, y = label_html)) +
    geom_tile(colour = "whitesmoke", fill = "white", size=0.5) +
    geom_point(aes(color = presence), size = 4) +
    scale_fill_manual(values = c(`TRUE` = "#08519C", `FALSE` = "#F0F0F0")) +
    scale_x_discrete(position = "top") +
    theme_minimal() +
    theme(axis.text.y = element_markdown(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank())
```

#### Pooling OTU counts at different taxonomic levels

A common task is to repeat OTU-level analyses at different taxonomic levels to
determine [which resolution is optimal for answering your questions](doi.org/doi/10.1128/mbio.03161-21). 
You'll need a shared file, generated from clustering sequences into OTUs with
mothur, and a corresponding taxonomy file.
Take a look at the [mothur documentation](https://mothur.org/wiki/) for info on
generating these files and performing microbiome analyses.

```{r pool_genus}
tax_dat <- read_tax(system.file("extdata", "test.taxonomy",
                                package = "schtools"))
shared_dat <- readr::read_tsv(system.file("extdata", "test.shared",
                                          package = "schtools"))
pool_taxon_counts(shared_dat, tax_dat, genus)
```

In this example, `pool_taxon_counts()` pools the OTU counts in the shared file 
at the genus level and returns new shared and taxonomy data frames.
You can do this for any taxonomic level in your taxonomy data frame.

```{r pool_phylum}
pool_taxon_counts(shared_dat, tax_dat, phylum)
```


### Distance files

If you have a distance file saved as a phylip-formatted lower triangle matrix
from mothur's [`dist.seqs`](https://mothur.org/wiki/dist.seqs/) command, 
you can read it into R with `read_dist()`:
 
```{r read_dist}
dist_filepath <- system.file("extdata",
                             "sample.final.thetayc.0.03.lt.ave.dist",
                             package = "schtools")
dist_tbl <- read_dist(dist_filepath)
head(dist_tbl)
```


## R Markdown helpers for scientific writing

When writing scientific papers with R Markdown, we often find ourselves 
using the same knitr chunk options and miscellaneous helper functions.
To use our favorite options like `eval=TRUE`, `echo=FALSE`, and others,
run `set_knitr_opts()` in the first chunk of your R Markdown document:


````markdown
`r ''````{r, include = FALSE}
set_knitr_opts()
```
````

This also sets the inline hook to our custom `inline_hook()` function, which
automatically formats numbers in a human-readable way and inserts an Oxford
comma into lists when needed.

### Who doesn't love an Oxford comma?

When writing with R Markdown, you may wish to insert a list or vector 
inline and correctly format it with an Oxford comma. 
`inline_hook()` uses `paste_oxford_list()` to help you do just that!

```{r oxford}
animals <- c('cats', 'dogs', 'fish')
```

Insert the string as inline code with `` `r ` ``:

> `` `r "\u0060r animals\u0060"` `` are the most common pets.

Rendered output:

> `r paste_oxford_list(animals)` are the most common pets.


### Human-readable numbers

`inline_hook()` uses `format_numbers()` under the hood to automatically format
numbers to a human-readable format, rather than display in scientific notation.

> The numbers `` `r "\u0060r c(1e-04, 1e-05, 1e-06)\u0060"` `` are very precise,
> while `` `r "\u0060r c(1e04, 1e05, 1e06)\u0060"` `` are very large.

Rendered output:

> The numbers `r inline_hook(c(1e-04, 1e-05, 1e-06))` are very precise.
> while `r inline_hook(c(1e04, 1e05, 1e06))` are very large.

