---
title: "Introduction to schtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to schtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

```{r setup}
library(schtools)
```

## Parsing mothur output files

### Taxonomy files

```{r read_tax}
tax_filepath <- system.file("extdata", "test.taxonomy", package = "schtools")
tax_dat <- read_tax(tax_filepath)
head(tax_dat)
```

The column `label_html` provides html that correctly italicizes the genus name
but not the OTU label, which can be used with `ggtext::element_markdown()`
to make nice plots:

```{r}
library(dplyr)
library(ggplot2)
library(ggtext)
library(tidyr)
set.seed(20220427)
pick_random_bool <- function(n = nrow(tax_dat)) {
        sample(c(TRUE, FALSE), n, replace = TRUE)
    }
tax_dat %>% 
    select(label_html) %>% 
    mutate(sample1 = pick_random_bool(),
           sample2 = pick_random_bool(),
           sample3 = pick_random_bool() 
           ) %>% 
    pivot_longer(starts_with('sample'), 
                 names_to = 'sample', values_to = 'presence') %>% 
    ggplot(aes(x = sample, y = label_html, fill = presence)) +
    geom_tile() +
    scale_fill_manual(values = c(`TRUE` = "#08519C", `FALSE` = "#F0F0F0")) +
    scale_x_discrete(position = "top") +
    theme_classic() +
    theme(axis.text.y = element_markdown(),
          axis.title.y = element_blank(),
          axis.title.x = element_blank())
```



### Phylip-formatted distance file as a tibble
 
```{r read_dist}
dist_filepath <- system.file("extdata",
                             "sample.final.thetayc.0.03.lt.ave.dist",
                             package = "schtools")
dist_tbl <- read_dist(dist_filepath)
head(dist_tbl)
```


## Miscellaneous helpers

When writing scientific papers with R Markdown, we often find ourselves 
using the same knitr chunk options and miscellaneous helper functions.
To use our favorite options like `eval=TRUE`, `echo=FALSE`, and others,
run `set_knitr_opts()` in the first chunk of your R Markdown document:


````markdown
`r ''````{r, include = FALSE}
set_knitr_opts()
```
````

This also sets the inline hook to our custom `inline_hook()` function, which
automatically formats numbers in a human-readable way and inserts an Oxford
comma into lists when needed.

### Who doesn't love an Oxford comma?

When writing with R Markdown, you may wish to insert a list or vector 
inline and correctly format it with an Oxford comma. 
`inline_hook()` uses `paste_oxford_list()` to help you do just that!

```{r oxford}
animals <- c('cats', 'dogs', 'fish')
```

Insert the string as inline code with `` `r ` ``:

> `` `r "\u0060r animals\u0060"` `` are the most common pets.

Rendered output:

> `r paste_oxford_list(animals)` are the most common pets.


### Human-readable numbers

`inline_hook()` uses `format_numbers()` under the hood to automatically format
numbers to a human-readable format, rather than display in scientific notation.

> The numbers `` `r "\u0060r c(1e-04, 1e-05, 1e-06)\u0060"` `` are very precise,
> while `` `r "\u0060r c(1e04, 1e05, 1e06)\u0060"` `` are very large.

Rendered output:

> The numbers `r inline_hook(c(1e-04, 1e-05, 1e-06))` are very precise.
> while `r inline_hook(c(1e04, 1e05, 1e06))` are very large.

