---
title: "Introduction to mothuR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to mothuR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mothuR)
```

## Parsing mothur output files

### Phylip-formatted distance file as a tibble
 
```{r read_dist}
dist_filepath <- system.file("extdata",
                             "sample.final.thetayc.0.03.lt.ave.dist",
                             package = "mothuR")
dist_tbl <- read_dist(dist_filepath)
head(dist_tbl)
```

## Miscellaneous helpers

When writing with R Markdown, you may wish to insert a list or vector 
inline and correctly format it with an Oxford comma. 
We provide `paste_oxford_list()` to help you do just that. 

```{r oxford}
animals <- c('cats', 'dogs', 'fish')
```

Insert the string as inline code with `` `r ` ``:

> `` `r "\u0060r paste_oxford_list(animals)\u0060"` `` are the most common pets.

Rendered output:

> `r paste_oxford_list(animals)` are the most common pets.
