---
title: "Introduction to schtools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to schtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

```{r deps}
library(schtools)
library(tidyverse)
```

## Handling mothur data

### Calculate relative abundances

You can read a shared file and calculate relative abundances with `calc_relabun()`:

```{r calc_relabun}
shared_dat <- read_tsv(system.file("extdata", "test.shared", 
                                   package = "schtools"))
relabun_dat <- shared_dat %>% calc_relabun()
head(relabun_dat)
```

`calc_relabun()` returns the data frame in long format. 
You can use `tidyr::pivot_wider()` to convert it to wide format:

```{r pivot_wider}
wide_dat <- relabun_dat %>% 
    pivot_wider(names_from = "otu", values_from = "rel_abun")
head(wide_dat)
```

You can see that the relative abundances for each sample sum to 1:

```{r sum1}
wide_dat %>% 
    select(starts_with('Otu')) %>% 
    rowSums()
```


### Taxonomy files

mothur formats taxonomy files as tab-separated values (tsv). 
You can use `read_tax()` to parse the taxonomy data and create separate columns
for each taxonomic level.

```{r read_tax}
tax_dat <- read_tax(system.file("extdata", "test.taxonomy", 
                                package = "schtools"))
head(tax_dat)
```

The column `label_html` provides html that correctly italicizes the genus name
without italicizing the OTU label. This can be used with
`ggtext::element_markdown()` to make nice plots:

```{r italic-genus}
library(ggtext)
set.seed(20220427)

relabun_dat %>%
    mutate(sample_num = str_remove(sample, 'p') %>% as.integer(),
           treatment = case_when(sample_num %% 2 == 1 ~ 'A', 
                                                 TRUE ~ 'B')
           ) %>% 
    inner_join(tax_dat, by = "otu") %>%
    ggplot(aes(x = rel_abun, y = label_html, color = treatment)) +
    geom_jitter(alpha = 0.7, height = 0.2) +
    labs(x = 'Relative abundance', y = '') +
    theme_minimal() +
    theme(axis.text.y = element_markdown())
```

#### Pooling OTU counts at different taxonomic levels

A common task is to repeat OTU-level analyses at different taxonomic levels to
determine [which resolution is optimal for answering your questions](https://doi.org/10.1128/mbio.03161-21). 
You'll need a shared file, generated from clustering sequences into OTUs with
mothur, and a corresponding taxonomy file.
Take a look at the [mothur documentation](https://mothur.org/wiki/) for info on
generating these files and performing microbiome analyses.

In this example, `pool_taxon_counts()` pools the OTU counts in the shared file 
at the genus level and returns new shared and taxonomy data frames.

```{r pool_genus}
tax_dat <- read_tax(system.file("extdata", "test.taxonomy",
                                package = "schtools"))
shared_dat <- readr::read_tsv(system.file("extdata", "test.shared",
                                          package = "schtools"))
pool_taxon_counts(shared_dat, tax_dat, genus)
```

You can do this for any taxonomic level in your taxonomy data frame.

```{r pool_phylum}
pool_taxon_counts(shared_dat, tax_dat, phylum)
```


### Distance files

If you have a distance file saved as a phylip-formatted lower triangle matrix
from mothur's [`dist.seqs`](https://mothur.org/wiki/dist.seqs/) command, 
you can read it into R with `read_dist()`:
 
```{r read_dist}
dist_filepath <- system.file("extdata",
                             "sample.final.thetayc.0.03.lt.ave.dist",
                             package = "schtools")
dist_tbl <- read_dist(dist_filepath)
head(dist_tbl)
```


## R Markdown helpers for scientific writing

When writing scientific papers with R Markdown, we often find ourselves 
using the same knitr chunk options and miscellaneous helper functions.
To use our favorite options like `eval=TRUE`, `echo=FALSE`, and others,
run `set_knitr_opts()` in the first chunk of your R Markdown document:


````markdown
`r ''````{r, include = FALSE}
set_knitr_opts()
```
````

This also sets the inline hook to our custom `inline_hook()` function, which
automatically formats numbers in a human-readable way and inserts an Oxford
comma into lists when needed.

### Who doesn't love an Oxford comma?

When writing with R Markdown, you may wish to insert a list or vector 
inline and correctly format it with an Oxford comma. 
`inline_hook()` uses `paste_oxford_list()` to help you do just that!

```{r oxford}
animals <- c('cats', 'dogs', 'fish')
```

Insert the string as inline code with `` `r ` ``:

> `` `r "\u0060r animals\u0060"` `` are the most common pets.

Rendered output:

> `r paste_oxford_list(animals)` are the most common pets.


### Human-readable numbers

`inline_hook()` uses `format_numbers()` under the hood to automatically format
numbers to a human-readable format, rather than display in scientific notation.

> The numbers `` `r "\u0060r c(1e-04, 1e-05, 1e-06)\u0060"` `` are very precise,
> while `` `r "\u0060r c(1e04, 1e05, 1e06)\u0060"` `` are very large.

Rendered output:

> The numbers `r inline_hook(c(1e-04, 1e-05, 1e-06))` are very precise.
> while `r inline_hook(c(1e04, 1e05, 1e06))` are very large.

